---
title: "Impact of Covid-19 on the U.S. Stock Market, by Phases and Sectors"
author: "ECON6082 Data Visualization Final Report"
date: "Feb 14,2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


| **Raissa Yeung **

------------------------------------------------------------------------

This Rmarkdown file mainly includes the code section and visualization graphs of the ECON 6082 final report. Please check the word document for full information.

## **Research questions**

Since the Covid-19 erupted, stock markets have been severely affected. We studied the representative S&P 500 index from the US market in **3 phases:** **(1)crash (2)recovery and (3)stable** identified according to the pandemic timeline. We aim to analyze how the Covid-19 changed the investment favor of investors based on the following dimensions:

1.  The linearity between stock market and Covid-19.

2.  How did the Covid-19 change the investment proportion in **different sectors** and in the **3 phases**? By comparing the US stock market **before and after the pandemic**.

3.  What sectors and firms are most **affected** or **defensive** during the Covid-19?

4.  What can we find by comparing the **correlation among sectors**?

## **Datasets**

The covid dataset is extracted from <https://covid.ourworldindata.org/data/owid-covid-data.csv> includes 160237 observations which shows worldwide Covid-19 cases starting from 24/2/2020 and updated in real time. The SP500 index from the R package `tidyquant` contained 517 observations showing the price, trade volume and sectors.

## Setup

```{r,warning = FALSE,message = FALSE}

# clear the environment
rm(list=ls())

setwd("~/Downloads/ECON6082 Data Visualization/final_project")

# load the packages
library(pacman)
p_load("ggplot2", "tidyverse", "knitr", "viridis", "hrbrthemes", "devtools", "readr","tidyquant","patchwork","gganimate","treemapify","corrplot",'reshape2',"sf","maptools","tigris","ggthemes","memisc")
```

```{r import data,warning = FALSE,message = FALSE}
# import covid data
covid <- read.csv("https://covid.ourworldindata.org/data/owid-covid-data.csv")

# get covid world total cases by date
covid %>% 
  filter(location=="World")  %>% 
  dplyr::select(date,total_cases,new_cases) %>% 
  mutate(date = as.Date(date)) ->data1


# get stock price
tq_get("^GSPC",get="stock.prices") %>%   # ticker"^GSPC"for SP500 Index
  dplyr::select(symbol,date,volume,close,adjusted) %>% 
  filter(date>=min(data1$date))->stock


# Join two data by date
data1 <- inner_join(data1,stock, by = "date")

```

### World Map

-   Distribution of the Covid-19 and Death rate

-   Seriousness of the pandemic

As shown on the map, Covid-19 is a worldwide pandemic. The color density represents the severeness of cases with darkest blue marks the highest. The red dots indicate the total death cases, the number is represented by the circle size.

```{r map, warning = FALSE,message = FALSE}

covid %>% 
  filter(date=="2021-12-08") %>% 
  group_by(iso_code) %>% 
  arrange(desc(date)) %>%
  slice(1:1) %>% 
  ungroup() %>% 
  dplyr::select(iso_code,location,date,total_cases,total_vaccinations,total_deaths,population)-> cov

# load the map boundries
data(wrld_simpl) # world boundaries
st_as_sf(wrld_simpl)-> world 
# join world
world <- geo_join(world,cov,"ISO3","iso_code")


# plot world total cases

pdf(file="worldmap.pdf",width=16, height=10)
ggplot(data = world)+
  geom_sf(aes(fill = total_cases), color="#b37400", size=0.1,alpha = 0.7) +
  scale_fill_gradient(low="#94e5ff",high="#498fd4",na.value="grey")+
  # scale_fill_distiller(palette = "Spectral")+
   xlab(NULL) +
  ylab(NULL)+
  geom_text(aes(x=LON, y=LAT+2, label=ISO3),
      color = "white", check_overlap = TRUE, fontface = "bold",size=3) +
  geom_point(data=world, aes(x=LON, y=LAT-2, size=total_deaths),
             alpha=0.7,color="#dd2500") +
  theme(panel.grid = element_blank(),
        plot.margin=unit(c(1,1,1,1), "cm"),
        plot.title = element_text(size=20,face="bold",color="white"),
        plot.subtitle = element_text(size=10,face="bold",color="white"),
        panel.background =element_rect(fill = "#384048"),
        plot.background=element_rect(fill = "#384048"),
        legend.background = element_rect(fill = "#384048"),
        legend.key = element_rect( fill = "#384048"),
        legend.text = element_text(size=10,color="white"),
        legend.title = element_text(size=13,color="white"),
        )+
  labs(title="World Total Cases", 
       subtitle = paste0("(",world$date[1],", ", length(unique(world$NAME)), " countries)"),
       fill="Total Cases",
       size="Total Deaths")
dev.off()
```

### linear regression

-   Investigate the linearity between X (i.e.Covid-19 cases) and Y (i.e. SP500 index)

-   If Covid do cause the significant impact on the US stock market.

In overall, the cumulative cases of Covid-19 and the SP500 index are positively correlated. The dark dots show the number of total cases of Covid-19 increases by time, the blue shadow indicates the relationship between the SP500 close price and the total cases of Covid-19. It can be seen that there is a linear relationship between them.

```{r reg, warning = FALSE,message = FALSE}


m1 <- lm(close ~ total_cases, data = data1)

mtable(m1)

pdf(file="regression.pdf",width=10, height=8)
ggplot(data = data1, aes(x = total_cases, y = close)) + 
  geom_smooth(method = "lm", #se=FALSE, 
              fill="#33CCCC",alpha=.25,formula = y ~ x) +
  geom_point() +
  scale_x_continuous()+
  theme_minimal()+
    theme(
      plot.title = element_text(colour = "black",face="bold", size = 20),
      plot.margin=unit(c(2,2,2,2), "cm"),
      plot.title.position = "plot",
      axis.text.x = element_text(colour = "#a9a9a9", size = 8),
      axis.text.y = element_text(colour = "#a9a9a9", size = 8),
 axis.title=element_text(size=12))+
  labs(title="Linear Relationship between Covid Total Cases and SP500 Index")+
  xlab("Total Cases")+
  ylab("SP500 Index")
dev.off()

```

### Sector Proportion：Trend Line Chart

In order to compare the impact of Covid-19, we identified 3 key phrases in terms of the line graph showing the market situation.

The pre-pandemic data archived between Jun 2019 - Dec 2019.

|                      |                     |
|----------------------|---------------------|
| 1st Phase(Outbreak)  | Jan.2020- Mar.2020  |
| 2nd Phase (Recovery) | Mar.2020 - Sep.2020 |
| 3rd Phase (Stable)   | Mar.2021 - Sep.2021 |

From the crash phase, all the sectors dropped sharply due to the black swan incident, among them the energy sector drop is the steepest. 

For the recovery phase, sectors took around a half year to recover. The market cap of IT, consumer discretionary(CD) and communication services(CS) even exceeded the pre-Covid period. Meanwhile, they experienced the fastest recovery speed among all. Utilities and health care sectors also performed well in their tiers, in which they recovered to their original market cap.

```{r line trend, warning = FALSE,message = FALSE}

# prepare data: date|sector|weight|rank

tq_index("SP500") %>%
  #slice(1:3) %>% 
  tq_get(get="stock.prices") %>% 
  filter(date>="2019-06-01") %>% 
  dplyr::select(symbol,company,weight,sector,shares_held,
         date,open,high,low,close,volume,adjusted) -> sp_500

#  [1] "symbol"         "company"        "identifier"    
#  [4] "sedol"          "weight"         "sector"        
#  [7] "shares_held"    "local_currency" "date"          
# [10] "open"           "high"           "low"           
# [13] "close"          "volume"         "adjusted"  

# group by date&sector, calculate the market value, add rank
sp_500 %>%   
  group_by(sector,date) %>% 
  summarize(mktcap_b=round(sum(shares_held*close/10^9),2)) %>% #calculate mkt cap in billions
  ungroup() %>% 
  group_by(date) %>% 
  mutate(rank=min_rank(mktcap_b),
         date=as.Date(date)) %>% 
  mutate(phase=
    if(date<="2020-01-01"){
      phase="pre"}
    else if (date>="2020-01-01"&date<"2020-03-27"){
      phase="crash"}
    else if(date>="2020-03-27"&date<"2020-10-01"){
      phase="recover"}
    else if (date>="2021-03-01"&date<"2021-10-01"){
      phase="stable"}
    else{phase=""})->data2

# line chart All time period (2019-09~now)

pdf(file="linechart_all.pdf",width=10, height=6)
ggplot()+
  geom_line(data=data2,aes(x=date,y=log(mktcap_b),group=rank(sector,mktcap_b),color=sector))+
  scale_color_manual(
    name="SP500 Stock",
    values=c("#004180","#006766", # Communication Services & consumer discretionary
             "#97c0e6",
             "#dc1f57", # Energy
              "#bebcff","#cad1d9","#b0d0d4",
             "#b52708", #Information Technology
             "#d9b8c8","#e8bebe","#dec4ab"))+
  scale_x_date(date_breaks = "4 month", date_labels =  "%Y-%m") +
  geom_ribbon(data=data1,
          aes(x=date,ymin=log(total_cases-new_cases*2)/4,ymax=log(total_cases)/4),fill="red",alpha=1)+
  theme_light()+
  theme(legend.key.height = unit(0.8,"cm"),
         plot.title = element_text(colour = "black",face="bold", size = 16),
      plot.margin=unit(c(1,1,1,1), "cm"),
      plot.title.position = "plot")+
  guides(color = guide_legend(override.aes = list(size = 4)))+
  #geom_line(data=data1,aes(x=date,y=total_cases),color="#A1B57D",alpha=0.7)+
  labs(title="Trends of Market Cap in Sections: All Time Periods",
       subtitle="  2019-06 ~ now",
    x="Date",y="Market Cap (Billions)")
dev.off()


# line chart multi-time period in panel

# pre-preiod: 2019 Sep-Dec
# stable:  2021 Mar-Sep

pdf(file="linechart_panel1.pdf",width=10, height=6)
ggplot()+
  geom_line(data=filter(data2,phase=="pre"|phase=="stable"),aes(x=date,y=log(mktcap_b),group=rank(sector,mktcap_b),color=sector))+
  facet_wrap(~phase, ncol=2,scales = "free_x")+
  scale_color_manual(
    name="SP500 Stock",
    values=c("#004180","#006766", # Communication Services & consumer discretionary
             "#97c0e6",
             "#dc1f57", # Energy
              "#bebcff","#cad1d9","#b0d0d4",
             "#b52708", #Information Technology
             "#d9b8c8","#e8bebe","#dec4ab"))+
  scale_x_date(date_breaks = "2 month", date_labels =  "%Y-%m") +
  theme_light()+
  theme(legend.key.height = unit(0.8,"cm"),
        plot.title = element_text(colour = "black",face="bold", size = 16),
      plot.margin=unit(c(1,1,1,1), "cm"),
      plot.title.position = "plot")+
  guides(color = guide_legend(override.aes = list(size = 4)))+
  #geom_line(data=data1,aes(x=date,y=total_cases),color="#A1B57D",alpha=0.7)+
  labs(title="Trends of Market Cap in Sections: Pre-covid vs. Stable Period",
       subtitle="  Before 2020-01 | 2021-03~2021-10 ",x="Date",y="Market Cap")
dev.off()


# stock market crash：2020 Jan-Mar
# recovery/defensive：2020 Mar-Sep

pdf(file="linechart_panel2.pdf",width=10, height=6)
ggplot()+
  geom_line(data=filter(data2,phase=="crash"|phase=="recover"),aes(x=date,y=log(mktcap_b),group=rank(sector,mktcap_b),color=sector))+
  facet_wrap(~phase, ncol=2,scales = "free_x")+
  scale_color_manual(
    name="SP500 Stock",
    values=c("#004180","#006766", # Communication Services & consumer discretionary
             "#97c0e6",
             "#dc1f57", # Energy
              "#bebcff","#cad1d9","#b0d0d4",
             "#b52708", #Information Technology
             "#d9b8c8","#e8bebe","#dec4ab"))+
  scale_x_date(date_breaks = "2 month", date_labels =  "%Y-%m") +
  theme_light()+
  theme(legend.key.height = unit(0.8,"cm"),
        plot.title = element_text(colour = "black",face="bold", size = 16),
      plot.margin=unit(c(1,1,1,1), "cm"),
      plot.title.position = "plot")+
  guides(color = guide_legend(override.aes = list(size = 4)))+
  labs(title="Trends of Market Cap in Sections: Crash Defensive vs. Recover",
       subtitle=" 2020-01~2020-03 | 2020-03~2020-10 ",x="Date",y="Market Cap")
dev.off()


```

### Sector Proportion：Dynamic Bar Chart

The dynamic bar chart shows the period of stock crash and recovery.

```{r Dynamic Bar Chart, warning = FALSE,message = FALSE}

# plot the dynamic bar chart (2020Jan-2020July)
data2 %>% 
  filter(date>="2020-01-01"&date<"2020-10-01")->data3
p2 <- ggplot(data3, aes(x=rank, y=mktcap_b, fill=sector)) +
    geom_col(width=0.5,alpha = 0.8) +
    geom_text(aes(y = -2, label = sector), vjust = 0.5, hjust = 1, show.legend = FALSE,size=4) +
    geom_text(aes(x=rank, y = mktcap_b+1, label = as.character(mktcap_b)), vjust = 0.5, hjust = 0, show.legend = F, col="grey18",size=8) +
    geom_text(x = 11.9 , y = -4, aes(label = as.character(data3$date)), size = 9, col = "#2F2F27") +
    coord_flip(clip = "off", expand = TRUE) +
    scale_x_reverse() +
    theme_ipsum(grid="X") +
    theme(panel.background = element_blank(),
          legend.position = "bottom",
          legend.title = element_blank(),
          plot.title = element_text(size = 24, face = "bold"),
          plot.title.position = "plot",
          plot.subtitle = element_text(size = 19,face="italic"),
          axis.text.y = element_blank()) +
    labs(y = NULL, #"Daily Sector Market Cap "
         x = NULL,
         title = "SP500 Market Cap During Covid-19 Pandemic in Billions, Sum by Sector",
         subtitle = "Time Period: Jan.2020~Oct.2020, Daily Frequency")  +
    scale_y_continuous(labels = NULL, limits = c(-25,110),breaks=seq(from=0,to=110,by=10))+
    scale_x_discrete()+
    #scale_fill_viridis(discrete = T, option = "viridis")+
    scale_fill_manual(values=c(
      "#f2f794", # 1
      "#c4e68a", #2
      "#7bedb0",#3
      "#2b498f", # 4
      "#305e8c",#5
      "#dff5f3",#6
      "#529ab3",#7
      "#8fcbcf",#8
      "#ed9015",#9
      "#e6bc8a",#10
      "#f2d666", #11
      "#a9a9a9"
    ))+
    transition_time(date)+ease_aes('cubic-in-out')

animate(p2, nframes = 300, fps = 25, width = 800, height = 600, renderer=gifski_renderer())

anim_save("mktcap_sector_1.gif")

```

### Ranking Bump Chart

**Energy**, **Materials**, **Real Estate and Utilities** are the most **volatile** sectors, however since the pandemic turning normalized, energy and materials sectors returned to their original states as the pre-pandemic period. 

IT is still staying strong and tough, which always rank first among the three phases. Therefore, it is the most favorable sector. CD and Financial sectors swapped positions, real estate and utilities changed positions as the former sector is more favorable for investors. The former is more favorable than the later one for investors.

```{r ranking bump,warning = FALSE,message = FALSE}

#arrange data:sector|date|rank|color
data2 %>%
  mutate(monthly=format(as.Date(date), "%Y-%m")) %>%
  group_by(sector,monthly) %>%
  summarize(mktcap_b=sum(mktcap_b)) %>%
  mutate(color=
           ifelse(sector=="Consumer Discretionary","#dd0049",
                  ifelse(sector=="Financials","#498fd4","white"))) %>%
  ungroup() %>%
  group_by(monthly) %>%
  mutate(rank=min_rank(mktcap_b),
         fake_date=as.Date(paste0(monthly,"-01",sep=""))) -> bump

pdf(file="bump.pdf",width=12, height=6)
ggplot(bump, aes(x = substr(fake_date,0,7), y = rank, group = sector)) +
  geom_line(aes(color = color)) +
  geom_point(aes(color = color),size=2) +
  expand_limits(y=c(0,12),x = c("2019-04", c("2023-06","2023-07","2023-08","2023-09","2023-10","2023-11","2023-12"))) +
  scale_color_identity() +
  scale_x_discrete(breaks=c("2019-06", "2020-01", "2020-03", "2020-09", "2021-03", "2022-02"))+
  geom_text(data = filter(bump, fake_date == "2019-06-01"), aes(x="2019-06",label = paste0(rank,"  "), color = color), hjust = 1, family = "sans",check_overlap = TRUE) +
  geom_text(data = filter(bump, fake_date == "2022-02-01"), aes(label = paste0("    ",sector), color = color), hjust = 0, family = "sans",check_overlap = TRUE) +
  annotate("text", x = "2020-02", y = 12, label = "Stock Crash", family = "sans",
  colour = "#e7e7e7", size = 5) +
  annotate("rect", xmin = "2020-01", xmax = "2020-03", ymin =0.5, ymax = 11.5, alpha = 0.3, fill = "#d3d3d3")+
    annotate("text", x = "2020-07", y = 12, label = "Recover", family = "sans",
  colour = "#e7e7e7", size = 5) +
  annotate("rect", xmin = "2020-04", xmax = "2020-10", ymin =0.5, ymax = 11.5, alpha = 0.3, fill = "#d3d3d3")+  
  annotate("text", x = "2021-06", y = 12, label = "Stable", family = "sans",
  colour = "#e7e7e7", size = 5) +
  annotate("rect", xmin = "2021-03", xmax = "2021-10", ymin =0.5, ymax = 11.5, alpha = 0.3, fill = "#d3d3d3")+

  theme(#axis.text.x = element_blank(),
        axis.text.x = element_text(colour = "#a9a9a9", size = 8),
        axis.ticks.x = element_line(colour = "#a9a9a9"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.margin=unit(c(1,1,1,1), "cm"),
        plot.title = element_text(colour = "white",size=22,margin = margin(0,19,10,100)),
        plot.subtitle=element_text(colour = "white",size=12,margin = margin(0,10,10,100)),
        plot.caption=element_text(colour = "white"),
        panel.grid=element_blank(),
        panel.background =element_rect(fill = "#384048"),
        plot.background=element_rect(fill = "#384048")
        ) +
  labs(x = NULL, y = NULL,
      title = "Few Sectors Changed Their Market Cap in the Covid Strike",
       subtitle = str_wrap("Ranking of market cap of SP500 sectors from 2019-2022 shows small cap industries are more volatile.", 100),
       caption = "Data: SP500")
dev.off()

```

## Tree Diagram

The tree diagram use sizes of rectangle to display each stock's market capital, color to indicate the increase and decrease of the stock price.

During the crash phase, the severely affected sectors are energy, financial and industrials as indicated by the color density. However, the stocks Tesla and REGN even gained a large market cap, which are 29.6% and 27.2% respectively.

Comparing the two periods, most of the stocks in pre-Covid period are growing. The color for TSLA is darkest as it sharply outperformed others and gained nearly 300% increase in market cap. The divergence of TSLA results in the overall tree map color lighter.

The color for the stable phase is obviously darker than the pre-Covid, the percentage change extent is more concentrated as there is no outlier. NVDA gained the largest increase(80.9%) ,TSLA(69.4%) and AMD(78%) also gained a relatively large increase.

```{r tree diagram, warning = FALSE,message = FALSE}

# tree diagram : crash
sp_500[which(sp_500$date=="2020-01-07"),] %>% 
  dplyr::select(symbol,company,weight,sector,shares_held,
         date,open,high,low,close,volume,adjusted) %>% 
  mutate(mktcap_b=shares_held*adjusted/10^6,
         date=as.Date(date))->temp

sp_500[which(sp_500$date=="2020-03-09"),] %>% 
  dplyr::select(symbol,company,weight,sector,shares_held,
         date,open,high,low,close,volume,adjusted) %>% 
  mutate(mktcap_b=shares_held*adjusted/10^6,
         date=as.Date(date),
         last_price=temp$adjusted,
         last_symbol=temp$symbol,
         last_date=temp$date) %>% 
  dplyr::select("symbol","last_symbol","company","sector","date","last_date","adjusted","mktcap_b","last_price") %>% mutate(change=round((adjusted/last_price-1)*100,4))->data4


pdf(file="treeDiagram_crash.pdf",width=20, height=20)
ggplot(data4, aes(
    area = mktcap_b, fill = change,
    subgroup = paste0(sector,"\n"), subgroup2 = paste0(symbol,"\n",change,"%",sep=""), label = "")) +
    geom_treemap(size = 0) +
    geom_treemap_text(place = 'middle', size = 8, min.size = 0, family = "sans") +
    geom_treemap_subgroup2_border(color = '#d3d3d3', size = 1) +
    geom_treemap_subgroup2_text(color = '#778899', place = 'topleft', size = 16, min.size = 3, family = "sans") +
    geom_treemap_subgroup_border(color = '#3b3b3b', size = 2) +
    geom_treemap_subgroup_text(color = '#272727', place = 'bottomright', size = 20, min.size = 3, family = "sans") +
  scale_fill_gradient2(high="#8b0000",mid="white",low="#008700")+
    #scale_fill_viridis_c(option = 'magma') +
    labs(title = 'Market Cap and Percentage Change(Period: Stock Crash)',
      subtitle = 'SP500 Stocks by Sector',     
      fill = 'Stock Price Percentage change')+ 
  theme(plot.title = element_text(size = 24, family = "sans", face = "bold"),
        plot.subtitle = element_text(size = 18, family = "sans"),
        plot.caption = element_text(size = 10, family = "sans"),
        legend.position = "bottom",
        legend.title=element_text(size=8, family = "sans"), 
        legend.text=element_text(size=8, family = "sans"))
ggsave("treeDiagram_crash.png", width = 15, height = 15, bg = 'transparent')
dev.off()


# tree diagram : pre
sp_500[which(sp_500$date=="2020-06-01"),] %>% 
  dplyr::select(symbol,company,weight,sector,shares_held,
         date,open,high,low,close,volume,adjusted) %>% 
  mutate(mktcap_b=shares_held*adjusted/10^6,
         date=as.Date(date))->temp

sp_500[which(sp_500$date=="2020-12-31"),] %>% 
  dplyr::select(symbol,company,weight,sector,shares_held,
         date,open,high,low,close,volume,adjusted) %>% 
  mutate(mktcap_b=shares_held*adjusted/10^6,
         date=as.Date(date),
         last_price=temp$adjusted,
         last_symbol=temp$symbol,
         last_date=temp$date) %>% 
  dplyr::select("symbol","last_symbol","company","sector","date","last_date","adjusted","mktcap_b","last_price") %>% mutate(change=round((adjusted/last_price-1)*100,4))->data4

pdf(file="treeDiagram_pre.pdf",width=20, height=20)
ggplot(data4, aes(
    area = mktcap_b, fill = change,
    subgroup = paste0(sector,"\n"), subgroup2 = paste0(symbol,"\n",change,"%",sep=""), label = "")) +
    geom_treemap(size = 0) +
    geom_treemap_text(place = 'middle', size = 8, min.size = 0, family = "sans") +
    geom_treemap_subgroup2_border(color = '#d3d3d3', size = 1) +
    geom_treemap_subgroup2_text(color = '#778899', place = 'topleft', size = 16, min.size = 3, family = "sans") +
    geom_treemap_subgroup_border(color = '#3b3b3b', size = 2) +
    geom_treemap_subgroup_text(color = '#272727', place = 'bottomright', size = 20, min.size = 3, family = "sans") +
  scale_fill_gradient2(high="#8b0000",mid="white",low="#008700")+
    #scale_fill_viridis_c(option = 'magma') +
    labs(title = 'Market Cap and Percentage Change(Period: Pre-Covid)',
      subtitle = 'SP500 Stocks by Sector',     
      fill = 'Stock Price Percentage change')+ 
  theme(plot.title = element_text(size = 24, family = "sans", face = "bold"),
        plot.subtitle = element_text(size = 18, family = "sans"),
        plot.caption = element_text(size = 10, family = "sans"),
        legend.position = "bottom",
        legend.title=element_text(size=8, family = "sans"), 
        legend.text=element_text(size=8, family = "sans"))
ggsave("treeDiagram_pre.png", width = 15, height = 15, bg = 'transparent')
dev.off()



# tree diagram : stable
sp_500[which(sp_500$date=="2021-06-01"),] %>% 
  dplyr::select(symbol,company,weight,sector,shares_held,
         date,open,high,low,close,volume,adjusted) %>% 
  mutate(mktcap_b=shares_held*adjusted/10^6,
         date=as.Date(date))->temp

sp_500[which(sp_500$date=="2021-12-31"),] %>% 
  dplyr::select(symbol,company,weight,sector,shares_held,
         date,open,high,low,close,volume,adjusted) %>% 
  mutate(mktcap_b=shares_held*adjusted/10^6,
         date=as.Date(date),
         last_price=temp$adjusted,
         last_symbol=temp$symbol,
         last_date=temp$date) %>% 
  dplyr::select("symbol","last_symbol","company","sector","date","last_date","adjusted","mktcap_b","last_price") %>% mutate(change=round((adjusted/last_price-1)*100,4))->data4

pdf(file="treeDiagram_stable.pdf",width=20, height=20)
ggplot(data4, aes(
    area = mktcap_b, fill = change,
    subgroup = paste0(sector,"\n"), subgroup2 = paste0(symbol,"\n",change,"%",sep=""), label = "")) +
    geom_treemap(size = 0) +
    geom_treemap_text(place = 'middle', size = 8, min.size = 0, family = "sans") +
    geom_treemap_subgroup2_border(color = '#d3d3d3', size = 1) +
    geom_treemap_subgroup2_text(color = '#778899', place = 'topleft', size = 16, min.size = 3, family = "sans") +
    geom_treemap_subgroup_border(color = '#3b3b3b', size = 2) +
    geom_treemap_subgroup_text(color = '#272727', place = 'bottomright', size = 20, min.size = 3, family = "sans") +
  scale_fill_gradient2(high="#8b0000",mid="white",low="#008700")+
    #scale_fill_viridis_c(option = 'magma') +
    labs(title = 'Market Cap and Percentage Change(Period: Stable)',
      subtitle = 'SP500 Stocks by Sector',     
      fill = 'Stock Price Percentage change')+ 
  theme(plot.title = element_text(size = 24, family = "sans", face = "bold"),
        plot.subtitle = element_text(size = 18, family = "sans"),
        plot.caption = element_text(size = 10, family = "sans"),
        legend.position = "bottom",
        legend.title=element_text(size=8, family = "sans"), 
        legend.text=element_text(size=8, family = "sans"))
#ggsave("treeDiagram_stable.png", width = 15, height = 15, bg = 'transparent')
dev.off()

```

## Correlogram

The correlation among sectors are highly correlated, ranging from the highest 0.92 to the lowest 0.43. In order to implement hedging strategies, investors may better invest sectors with a low correlation which can help them to reduce risk in some unstable period or unexpected shocks.

```{r correlation,warning = FALSE,message = FALSE}

# sector change rate was calculated as the total sector market cap's change rate.

# calculate the daily mktcap(billions) change(%)
data2 %>% 
  mutate(pre_date=date-1) %>% 
  filter(date>="2020-01-03")->data5

temp <- inner_join(data2,data5, by =c("date"="pre_date","sector"="sector"))
colnames(temp) <- c("sector","pre_date","pre_mktcap_b","pre_rank","pre_period","date","mktcap_b","rank","period")
temp %>% mutate(change=round((mktcap_b/pre_mktcap_b-1)*100,2)) ->data5

# prepare data in columns
data5 <- dcast(data5,date~sector)[,2:12]
colnames(data5) <- c("TELS","COND","CONS","ENRG","FIN","HLTH","INDU","INFT","MATR","REAS","UTIL")

# reorder the columns according to mean(mktcap)
data5 <- data5[,c(8,6,5,1,2,7,3,4,11,10,9)] 
  

# graph the plot

pdf(file="correlogram.pdf",width=8, height=6)

corrplot(cor(data5), method="number",type="upper",
        order = 'original',
         tl.col = "grey",number.font=1,
         tl.cex=1,mar=c(5,5,5,5),
         title="Correlation Between Sectors",tl.pos="l",addgrid.col="grey",
        col = COL2('RdBu', 11)) %>% 
  corrRect(index = c(1,2,8,11))
corrplot(cor(data5), 
         method = 'circle', 
         order = 'original',
         number.font=1,addgrid.col="grey",
         tl.cex=1,
         mar=c(5,5,5,5),tl.pos="n",cl.pos='n',
        add=T,type = 'lower', diag = FALSE) %>% 
  corrRect(index = c(1,2,8,11))

dev.off()

```

## Reference

*Market Phases Led by COVID Trends: What Do They Tell Us?(24 Sep 2021), HKTDC Research. Retrieved from <https://research.hktdc.com/en/article/ODYxNDg2MjM4>*

*Mathieu, E., Ritchie, H., Ortiz-Ospina, E. et al. A global database of COVID-19 vaccinations. Nat Hum Behav (2021) Retrieved from[**https://ourworldindata.org/covid-vaccinations**](https://ourworldindata.org/covid-vaccinations)*
